Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-08 paper.tex
Start writing "Narrative Interface" section.

\begin{abstract}
Project Nell%

is tablet software to teach reading and writing to
children far from educational infrastructure.  Nell is a modular
narrative system with ``a low floor and no ceilings'': the design
scales from teaching letter shapes to programming the system itself.

Nell is the successor to the Sugar educational software developed by
One Laptop per Child for their XO-1 laptop.  Nell targets the XO-3
tablet, a sub-\$100 solar-powered tablet for the developing world.
Experience with Sugar drove design choices for Nell.
\end{abstract}

@@

As Rahel grows, the book asks her to trace not just letters, but whole
words.   The books responses are written on the screen as it writes
them, and eventually she doesn't need to leave the sound on all the
time to read what the Princess is saying.  Soon Rahel can write
complete sentences in her special book, and the Princess will respond
sometimes.  New stories teach her about music (by playing certain
tunes she unlocks a dungeon door) and programming with blocks (a
not-very-bright turtle needs her to tell him how to draw different
shapes).  Rahel writes her own stories, which she shares with her
far-away friends.

An older Rahel learns that the block language she used to talk with
the turtle is actually used to write all the software running inside
her special book.  The stories Rahel writes can now include new
functionality that make her book even more exciting for the kids
just now opening their green meta books.

@@

We believe the combination of these four key ideas provides a novel
experience which addresses some of the shortcomings of earlier
learning environments for children.  In this section we will discuss
our motivating core concepts, and in the next we will describe how
they are implemented in Nell.

@@

\subsection{Narrative}
% excerpt of story model
\subsection{Personal}
% customization, renaming characters
\subsection{Direct interaction}
% handwriting as UI?
% motor skills
% photo of XO-3?
\subsection{Growable (``no ceiling'')}
% it has a ``low floor and no ceiling'' to allow a child to master more
% and more of the system as she grows.

@@

\section{Related work}\label{sec:related}
% Nell focused on Literacy
% olpc context here?
\cite{chang:tinkrbook, flores:uruguay, idb:peru}
%context: english literacy, far from teachers & schools
\subsection{The Sugar Learning Environment}
% Sugar
%% customization
%% path through the educational material?
\subsection{Narrative computing}
% Reidl
\subsection{Storytelling}
% TinkRBooks

@@

>>>>

\begin{abstract}
Project Nell is tablet software to teach reading and writing to
children far from educational infrastructure.  Nell is a modular
narrative system with ``a low floor and no ceilings'': the design
scales from teaching letter shapes to programming the system itself.

Nell is the successor to the Sugar educational software developed by
One Laptop per Child for their XO-1 laptop.  Nell targets the XO-3
tablet, a sub-\$100 solar-powered tablet for the developing world.
%Experience with Sugar drove design choices for Nell.

\textbf{XXX REWRITE THE ABSTRACT XXX}
\end{abstract}

@@

As Rahel grows, the book asks her to trace not just letters, but whole
words.  The book's responses are written on the screen as it speaks
them, and eventually she doesn't need to leave the sound on all the
time.  Soon Rahel can write complete sentences in her special book,
and the Princess will respond sometimes.  New stories teach her about
music (by playing certain tunes she unlocks a dungeon door) and
programming with blocks (a not-very-bright turtle needs Princess Rahel's
help to draw different shapes).  Rahel writes her own stories,
which she shares with her friends.

An older Rahel learns that the block language she used to talk with
the turtle is also used to write all the software running inside
her special book.  The stories Rahel writes can now include new
functionality that make her book even more exciting for the kids
just now opening their green meta books.

@@

We believe the combination of these four key ideas provides a novel
experience which addresses some of the shortcomings of earlier
learning environments for children.  In this section we will discuss
our motivating concepts, and in the next we will describe how
they are implemented in Nell.

@@

\subsection{Narrative Interface}

Experience with our previous educational software
system~\cite{flores:uruguay,idb:peru} has shown gratifying uptake and
use by children.  But we are often disappointed that some of our
favorite pedagogical tools included in the systems were not being
widely used.  Enthusiastic teachers would spark intensive use in
certain classrooms, but children had difficulty finding their way into
the material without strong guidance.

Children (like all humans) are hard-wired for
stories~\cite{boyd:stories}.  Nell uses a
\textit{Narrative Interface}~\cite{bizzocchi:narrative,don:narrative,laurel:computers} 
to guide the child user through the pedagogical capabilities of the system.

% excerpt of story model
% not always a story -- kid can draw on their own and return to the
% story; Nell will observe the free play and offer suggestions.
\subsection{Personal}
% customization, renaming characters
\subsection{Direct interaction}
% handwriting as UI?
% motor skills
% photo of XO-3?
\subsection{Growable (``no ceiling'')}
% it has a ``low floor and no ceiling'' to allow a child to master more
% and more of the system as she grows.

@@

\section{Related work}\label{sec:related}
% Nell focused on Literacy
% olpc context here?
\cite{chang:tinkrbook}
%context: english literacy, far from teachers & schools
\subsection{The Sugar Learning Environment}
% Sugar
%% customization
%% path through the educational material?
\subsection{Narrative computing}
% Reidl
\subsection{Storytelling}
% TinkRBooks

@@

\section{Acknowledgments}
Nell's design evolved from experience with the Sugar system,
architected by Walter Bender, and through stimulating discussions with
Chris Ball and Angela Chang.  Michael Stone also contributed to an
early implementation.
